


With each new addition of a variable, we see that the adjusted R-squared is increasing indicating that the model fit is improving. However, the fit statistic is still quite low. Note that the 25 to 39 years age category has insignificant p-values. 

# VIF Table to select variables 
```{r}
ezids::xkablevif(reg_3, wide=FALSE)
```

From, the VIF table, The VIF value for sleep time is 11.3, this is undesirable and thus will be dropped. Race is pretty above 5 for whites and other races, race as a variable will be dropped. 

## Logistic Regression  

# Model 1: HeartDisease~ Sex+ BMI
```{r}
# Change variables to factors 
heart$HeartDisease<-factor(heart$HeartDisease)
heart$Sex<-factor(heart$Sex)

# Logit Model 1
model_1<- glm(HeartDisease~ Sex+ BMI, data=heart, family='binomial')
#summary(model)

# Chi-square for Model 1
1-pchisq(186906 - 184556, df=(319794-319792))

# Exponential Coefficients  & Confidence Interval 
exp(cbind(OR = coef(model_1), confint(model_1)))

```

Calculating the chi-squared p-value of the difference between the null and residual deviance will tell us how "useful" the model is. Since, the chi-squared p-value of 0 is less than 0.05, we can conclude that the logit model with sex and BMI as predictors is good fit model for predicting the probability of getting heart disease. 

According to model 1, the log odds for getting heart disease increases by 1.64 for a man compared to a woman. While an increase in BMI increases the log odds for heart disease by 1.02. 

# Model 2: HeartDisease~ Sex+ BMI + PhysicalHealth + MentalHealth +  AgeCategory
```{r}

# Logit Model 2 
model_2<- glm(HeartDisease~ Sex+ BMI + PhysicalHealth + MentalHealth +  AgeCategory, data=heart, family='binomial')
#summary(model_2)

#Chi-squared test for model fit 
1-pchisq(186906 - 157118, df=(319794-319772))

# Exponential Coefficients  & Confidence Interval 
exp(cbind(OR = coef(model_2), confint(model_2)))
```

Just like before, all the coefficient p-values for our second model is significant and the chi-squared test is indicating that model 2 is also a good fit. This will be our preferred model for this section because it has a lower AIC score than model 1. 


did not have enough lifestyle variables, 
death variables: bevause of previous reference about women dying more 

# Cost Analysis 