---
title: "Key Risk Factors of Heart Disease in 2020"
author: "Team One"
date: April 4, 2022
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
#Introduction 

According to the Center for Disease Control and Prevention (CDC), cardiovascular or heart diseases account for 1 in every 4 deaths in the United States annually. The estimated costs to the nation’s economy amount to about $363 billion annually, which include hospitalizations, after services health care needs as well as lost productivity due to death and disability (SS Virani, 2021).  The CDC also states that given the lifestyle habits of an average American, about half of the population (47%) are reported to have “1 of the 3 key risk factors for heart disease: high blood pressure, high cholesterol, and smoking”. However, the risk factors for developing heart diseases are vast and diverse and tend to vary between most racial and ethnic groups in the United States. 

Therefore, there is an immediate need to detect the major risk factors early on, so that a wellness-oriented healthcare system can be instituted. Ideally, such a system would assess the intensity of the risk factors (compared to a baseline) and focus efforts on preventing the occurrence of developing heart diseases in the first place rather than minimizing second occurrences (Liau et al, 2010; Giampaoli et al, 2005 ). 

The beginnings of the current understanding of the interconnected role of risk factors was first discussed in the Framingham Heart Study. Started in 1948, the study’s original goal was “ to identify common factors or characteristics that contribute to cardiovascular disease” (Hong 2018). However, today the FHS has morphed into a multigenerational study gathering genetic information that help analyze family patterns of certain diseases, including cardiovascular ailments (National Heart, Lung and Blood Institute). Jumping off from this, current research in the field has focused on using precision medicine algorithms along with computational development technology to assess a patient’s likelihood of developing heart disease given certain lifestyle factors. Most of the studies have focused largely on Caucasian individuals as a sample and indicating that there is a gap in the usefulness of current risk assessment tools working for individuals of differenet races. 

In this paper, we aim to understand the different risk factors that affect the occurrence of heart diseases in the US population, so that we may be able to see some of the intermingling between different risk factors.   

#Description of Dataset

The original data set can be retrieved from the CDC’s Behavioral Risk Factor Surveillance System (BRFSS) annual health survey. The survey interviews over four hundred thousand individuals in the USA asking questions regarding the risk factors of heart disease. Conducted in 2020 (published on February 15, 2020), the original data set has been cleaned from its original 300 variables to 18 variables by Kamil Pytlak and can be accessed on Kaggle . 

*** Table with variable names****


```{r}
# Lets load the appropriate libraries
library(tidyverse)
library(dplyr)
library(funModeling)
library(ezids)
library(ggplot2)
library(ggcorrplot)
library(corrr)
library(corrplot)

# Load the Data 
heart <-read.csv("heart_2020_cleaned.csv")
nrow (heart)
```


# Descrpitive Statistics  
```{r}
status (heart)
```

From the table we can see that physical and mental health have a higher percentage of zeros. This means that just over 70% of the people in our data set are report physically sick in a 30 day period while about 60% of the respondents report feeling mentally unwell in that same time period.  

We can also see that there are no N/A values in our data set. This is not surprising given that our data was cleaned before being retrieved.  


```{r }
# Finding the mean of numerical variables grouped by heart disease
plot_num(heart)

heart %>%
  group_by(HeartDisease) %>%
  summarise_at(vars("BMI", 
                    "PhysicalHealth", 
                    "MentalHealth",
                    "SleepTime"), mean)
```

On average, people with heart disease have a slightly higher BMI than those who don't have heart disease. However, the biggest difference is the number of days that a respondent feels physically unwell. Respondents who have heart disease reported feeling physically unwell about 8 days per month, while healthy people felt unwell for only 3. Poor mental health in a 30 day period is slightly higher for people with heart disease, than for people without. 

```{r}
# Frequency distributions of all our categorical variables 
freq(heart)

# Changing Diabetes bordeline and pregnancy categories to Yes/No

heart$Diabetic<- replace (heart$Diabetic, heart$Diabetic== "No, borderline diabetes" , "Yes")
heart$Diabetic[heart$Diabetic== "Yes (during pregnancy)"] <- "No"  

#Checking if diabetic has been changed 

describe (heart$Diabetic)
freq(heart$Diabetic)

# Changing all the categorical variables to numerical dummies 

heart$HeartDisease<-ifelse(heart$HeartDisease=="Yes",1,0)
heart$Smoking<-ifelse(heart$Smoking=="Yes",1,0)
heart$AlcoholDrinking<-ifelse(heart$AlcoholDrinking=="Yes",1,0)
heart$Stroke<-ifelse(heart$Stroke=="Yes",1,0)
heart$DiffWalking<-ifelse(heart$DiffWalking=="Yes",1,0)
heart$PhysicalActivity<-ifelse(heart$PhysicalActivity=="Yes",1,0)
heart$KidneyDisease<-ifelse(heart$KidneyDisease=="Yes",1,0)
heart$SkinCancer<-ifelse(heart$SkinCancer=="Yes",1,0)
heart$Diabetic<-ifelse(heart$Diabetic=="Yes",1,0)
heart$Asthma<-ifelse(heart$Asthma=="Yes",1,0)
```

#Correlation Plots of All Variables
```{r}
 model.matrix(~0+., data=heart) %>% 
     cor(use="pairwise.complete.obs") %>% 
   ggcorrplot(show.diag = F, type="upper", lab=TRUE, lab_size=2)
```


## Gender and Lifestyle Factors

According to Noel Bairey Merz, MD, director of the Barbra Streisand Women’s Heart Center in the Smidt Heart Institute, women experience heart disease more severely than men.  Meaning that even though men are more likely to suffer heart attacks, women are more likely to die from a heart attack than a man. 

We started off by trying to see which gender tends to have more heart disease based on certain lifestyle choices.  From the table below, we can see that we have a lot more data for people without heart disease than with. However, even in the percentage table for those who do have heart disease, we see that there are more men with heart disease (about 5%) compared to women (3%). This is in line with our basic theory that men have more heart disease.  However, we are then forced to ask why do men have more heart disease in our sample? Are they doing more or less of a certain activity than their female counterparts? Discussions of the relevant variables are provided below. 


<<<<<<< HEAD
# Plots for Sex & Heart Disease 
=======
#Contingency Table for Sex & Heart Disease
>>>>>>> 25656f97d1b26d2c00b9ffe168f8f5f790b769a8

```{r}

contable = table(heart$heartDisease, heart$Sex)
xkabledply(contable, title="Contingency Table for heart Disease and Sex")
addmargins(contable)

prop <-prop.table(contable)
percent <- round(prop*100, 2)

print ("Contingency Table for heart Disease and Sex in Percentage")
print (percent)
barplot(with(heart, table(heartDisease, Sex)), main="heart Disease & Sex", beside=T, col=3:2)
```


<<<<<<< HEAD
In our exploratory analysis, we compared the different lifestyle variables between men and women and then further compared those same variables between those with and without heart disease. 
=======
However, we are then forced to ask why do men have more heart disease in our sample? Are they doing more or less of a certain activity than their female counterparts?

In our exploratory analysis, we compared the different lifestyle variables between men and women and then further compared those same variables between those with and without heart disease.
>>>>>>> 25656f97d1b26d2c00b9ffe168f8f5f790b769a8

##Body Mass Index (BMI)

# Is there a difference between the mean BMI of men and woman in the overall population? 

```{r}
hist(heart$BMI)
```

In the historgram, we can see that our data is approximately normal and given the Central Limit Theorem, if n>30, then we can assume normality for the sake of statistical tests.

After sub-setting the data between male and female, we conducted a Welch's t-test on the BMIs of both population subsets. 

```{r}
# Subset Male & Female with and without Heart Disease
female <- subset(heart, Sex== "Female")
male  <- subset (heart, Sex == "Male")

# T-test of BMI between Genders 
t.test(female$BMI, male$BMI)
```

The p-value is less than 5% significance level, allowing us to reject the null hypothesis and conclude that the overall average BMI is different between the male and female populations. From the mean values, we see that men have a mean BMI of 28.50 while women have a BMI of 28.16. A difference of 0.34.  

Next, we divided our data frame between men and women who do have heart disease. Once again assuming normality due to the large sample size, we conducted another Welch's T-test to look at the mean BMI difference. 

# Is there a difference in average BMI between the sexes with heart disease?

```{r}
# Subsetting genders based on respondents that have heart disease 
HD <- subset(heart, HeartDisease==1)
female_HD<- subset(female, HeartDisease==1)
male_HD<- subset(male, HeartDisease== 1)

#T-Test
t.test(female_HD$BMI, male_HD$BMI)
                  
 
```

From the T-test output above, we see that the p-value is 0.604, which is greater than 0.05. This means that we fail to reject the null and conclude that there is no significant difference in the average BMI between men and women who do have heart disease. 


##Physical Health 
# How many times in the last 30 days have you felt physically ill? 

Earlier, we saw that respondents, who have heart disease, reported feeling physically unwell about 8 days per month, while healthy people felt unwell for only 3. The box plot below shows us that women with heart disease generally report felling unwell more than men. 

```{r}
ggplot(heart, aes(x = factor(HeartDisease), y = PhysicalHealth))+
    geom_boxplot(aes(fill = Sex))+
    xlab('HeartDisease') + 
    scale_fill_brewer(palette = 'Set1', name = 'Sex')
```

```{r}
t.test(male_HD$PhysicalHealth,
female_HD$PhysicalHealth)
```

The t-test also confirms what the box plot is showing. Since the p-value is less than 5%, that there is a difference in the number of days that men and women feel unwell. On average, women reported feeling sick about 9 days while men reported just over 6 days. 


##Mental Health 
# How many times in the last 30 days have you felt physically ill? 
```{r}
ggplot(heart, aes(x = factor(HeartDisease), y = MentalHealth))+
    geom_boxplot(aes(fill = Sex))+
    xlab('HeartDisease') + 
    scale_fill_brewer(palette = 'Set1', name = 'Sex')

t.test  (male_HD$MentalHealth,
female_HD$MentalHealth)
```

The boxplot shows that women tend to report feeling mentally unwell more often then men.Since the p-value is less than 5%, indicating that there is a difference in the number of days that men and women feel mentally unwell. On average, women reported feeling mentally unwell about twice as many days as men.

##Sleep
#On average,how many hours of sleep do you get in a 24-hour period?
```{r}
ggplot(heart, aes(x = factor(HeartDisease), y = SleepTime))+
    geom_boxplot(aes(fill = Sex))+
    xlab('HeartDisease') + 
    scale_fill_brewer(palette = 'Set1', name = 'Sex')

t.test  (male_HD$SleepTime,
female_HD$SleepTime)
```

Men and women have statistically different sleep times. Men on average get about 35 more minutes of sleep a week than women. This may not sound like much but sleep effects are known to be cumulative. A study conducted by Robbins et al. in 2021 "examined a group of people who got the recommended seven to eight hours of sleep a night to those who got less (even just 30 minutes less), and exposed them to a norovirus". The study found that curtailing sleep below the recommended seven hours increases the risk for viral infection "nearly four times higher among those who are sleeping poorly or getting insufficient sleep, compared to those who are getting good rest" (Robbins et al, 2021). 

This lack of sleep could explain why we see that women in our data on average report feeling sick more often and could also be a contributing factors is why women are more likely to die from heart attacks than men (Noel Bairey Merz).

When looking at other lifestyle factors such as smoking or alcohol drinking, we see that there is no statistical difference that can be measured between the sexes and those that have heart disease. However, it should be noted that the survey is voluntary and there is no way to ensure that the respondents are being truthful. 

## Age, Race, Gen Health
```{r}
freq(HD$AgeCategory)
freq(HD$Race)
freq(HD$GenHealth)

```

When looking at the population make-up of respondents who do have heart, we see that a higher percentage of respondents are 60 and older, with 82% being of Caucasian descent. At the time of the interview they also said they felt their overall health is good. This may be a limiting factor for future research since, once again, the sample does not have many data points from non-Caucasian individuals.  We also see that most of our respondents with heart disease are older above 60 years of age, this is in-line with theory. 


# Liner Regression on Sex
```{r}
reg_1 <- lm(HeartDisease~Sex+BMI+ PhysicalHealth+MentalHealth, data=heart)
summary (reg_1)

reg_2 <- lm(HeartDisease~Sex+BMI+ PhysicalHealth+MentalHealth + SleepTime, data=heart)
summary (reg_2)

reg_3 <- lm(HeartDisease~Sex+BMI+ PhysicalHealth+MentalHealth + SleepTime +AgeCategory + Race, data=heart)
summary (reg_3)
```

Adjusted r-square is increasing so definitely model fit is improving but still not a good fit 


###########################################################################################

# heart Disease & Secondary Illnesses 
In this part we are Identifying whether heartDisease and secondary illnesses have correlation or not(are they dependent or independent?).

```{r}
# diseases <- data.frame (heart$heartDisease,
#                         heart$Stroke,
#                         heart$Diabetic,
#                         heart$Asthma, 
#                         heart$KidneyDisease,
#                         heart$SkinCancer)
diseases<-subset(heart,select = c(HeartDisease,Stroke,Diabetic,Asthma,KidneyDisease,SkinCancer))

# Rename columns

names(diseases) <- c ('heart_disease',
                      'stroke',
                      'diabetic',
                      'asthma',
                      'kidney_disease',
                      'skin_cancer')


```

# Corellation  Matrix between diseases
```{r}
model.matrix(~0+., data=diseases) %>% 
     cor(use="pairwise.complete.obs") %>% 
    ggcorrplot(show.diag = F, type="upper", lab=TRUE, lab_size=2)

```

Correlations not particularly strong 
heart disease and stroke seem correlated
Diabetes is correlated with heart disease and Kidney disease. Less with Stroke.
Skin cancer is not correlated with other  diseases.

# Determining Independence between Diseases
The Chi-Square test of independence is used to determine if there is a significant relationship between two nominal (categorical) variables(All the variables we have here are already categorical).  The frequency of each category for one nominal variable is compared across the categories of the second nominal variable.  The data can be displayed in a contingency table where each row represents a category for one variable and each column represents a category for the other variable. The null hypothesis for this test is that there is no relationship between primary disease heartDisease and other secondary illnesses.  The alternative hypothesis is that there is a relationship between heartDisease and secondary illnesses.

output:
the title of the test,
which variables have been used,
the test statistic,
the degrees of freedom and
the p-value of the test.

```{r}
hea_as=table(heart$heartDisease,heart$Asthma)
chi_1=chisq.test(hea_as)
chi_1

hea_dia=table(heart$heartDisease,heart$Diabetic)
chi_2=chisq.test(hea_dia)
chi_2

hea_kid=table(heart$heartDisease,heart$KidneyDisease)
chi_3=chisq.test(hea_kid)
chi_3

hea_skin=table(heart$heartDisease,heart$SkinCancer)
chi_4=chisq.test(hea_skin)
chi_4

hea_stroke=table(heart$heartDisease,heart$Stroke)
chi_5=chisq.test(hea_stroke)
chi_5
```

For all the chi^2 tests we performed above, the p-value<0.05. So all the above rejects null hypothesis. 
Every secondary illness (Diabetic,Asthma, Kidney disease and skin cancer) has a relationship with heart Disease. 


# Smokers Vs Drinkers 


```{r}
contable_smoke = table(heart_2020_cleaned$heartDisease, heart_2020_cleaned$Smoking)

prop <-prop.table(contable_smoke)
percent <- round(prop*100, 2)

xkabledply(percent, title="Contingency Table for heart Disease and Smoking in Percentage")
```

```{r}

contable_drink = table(heart_2020_cleaned$heartDisease, heart_2020_cleaned$AlcoholDrinking)

prop <-prop.table(contable_drink)
percent <- round(prop*100, 2)

xkabledply(percent, title="Contingency Table for heart Disease and Drinking in Percentage")

```

```{r chisq}

chisq.test(contable_smoke)

chisq.test(contable_drink)

```
```
For both the chi^2 tests we performed above, the p-value<0.05. So all the above rejects null hypothesis. 
Both drinking and smoking have a relationship with heart Disease. 